A fast unified model for parsing and sentence understanding

### 2016-03-16 writing/hard_stack_paper/paper.tex
Inputting CM/CP comments from today's meeting.

Of these, the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}. However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.  Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets. Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations. In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser. In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.

@@

\citet{socher2011semi} presents a version of the TreeRNN model which is capable of operating over unparsed inputs. However, this method is unsupervised, limiting its performance, and it requires an expensive search process at test time. Our model presents an alternative approach to combining interpretation and parsing that supports supervised training for the parsing component, and that is dramatically faster than any previous tree-structured model.

@@

The TreeLSTM composition function generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$. Our version is formulated as:
\begin{gather}
\colvec{5}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
    {\vec{g}}
= \colvec{5}
    {\sigma}
    {\sigma}
    {\sigma}
    {\sigma}
    {\text{tanh}}
\left(
W_{\text{iffo}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
\right) \label{eqn:lstm1}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}
$\sigma$ is the sigmoid activation function. The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}}
\end{equation}

@@

For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply this as part of the input. To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
\begin{equation}
\vec{p}_{\text{t}} = \text{softmax}(W_{\text{transition}}\vec{h}_{\text{tracking}})
\end{equation}
In the full SPINN model, we follow whichever transition (i.e., \shift~or \reduce) has a higher predicted probability in this distribution. This prediction function is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to Stanford PCFG Parser constituency parses included with SNLI.

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lcccc} 
    \toprule
Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
\midrule
\multicolumn{5}{c}{Previous non-NN results}\\
\midrule
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
\midrule
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{Our results}\\
\midrule
\textbf{300D LSTM RNN encoders}          & 3.0m                  & --                &   83.9      &   80.6       \\
\textbf{300D SPINN-PI-NT encoders}
                        & 3.4m                  & --                &   84.4      &   80.9       \\
\textbf{300D SPINN-PI encoders}
                        & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
\result{\textbf{300D SPINN encoders} }
                        & ??                  & 92?           &   ?    &   82?      \\          
% \result{300D SPINN-PI, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-PI, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
\midrule
\multicolumn{5}{c}{Other previous NN results}\\
\midrule
100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
                        & 252k               & --              &   85.3   &   83.5       \\
300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
                        & 1.9m               & --             &   92.0   &   86.1      \\
300D LSTMN with deep attention fusion \citep{cheng2016long}
                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy.} 
\end{table*}

@@

\question{What kind of error analysis would be interesting for this model?}

\result{[SB] Compare the types of errors in the different runs.}

>>>>

Of these, the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}. However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.  Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets. Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations. In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser. In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models. \todo{[SB] Weaken parsing claim.}

@@

\citet{socher2011semi} and \citet{socher2011parsing} present versions of the TreeRNN model which are capable of operating over unparsed inputs. However, these methods require an expensive search process at test time. Our model presents an alternative approach to combining interpretation and parsing that supports supervised training for the parsing component, and that is dramatically faster than any previous tree-structured model.

@@

The TreeLSTM composition function generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$. Our version is formulated as:
\begin{gather}
\colvec{5}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
    {\vec{g}}
= \colvec{5}
    {\sigma}
    {\sigma}
    {\sigma}
    {\sigma}
    {\text{tanh}}
\left(
W_{\text{comp}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
+ \vec{b}_{\text{comp}}
\right) \label{eqn:lstm1}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}
$\sigma$ is the sigmoid activation function. The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform the representation of each input word $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
\end{equation}

@@

For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply this as part of the input. To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
\begin{equation}
\vec{p}_{\text{t}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{tracking}} + \vec{b}_{\text{trans}})
\end{equation}
In the full SPINN model, we follow whichever transition (i.e., \shift~or \reduce) has a higher predicted probability in this distribution. This prediction function is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to Stanford PCFG Parser constituency parses included with SNLI.

@@

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lcccc} 
    \toprule
Model                   & Params.    & Trans. acc.  &   Train  &   Test \\
\midrule
\multicolumn{5}{c}{Previous non-NN results}\\
\midrule
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{Previous sentence encoder-based NN results}\\
\midrule
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{Our results}\\
\midrule
\textbf{300D LSTM RNN encoders}          & 3.0m                  & --                &   83.9      &   80.6       \\
\textbf{300D SPINN-PI-NT (parsed input, no ...) encoders}
                        & 3.4m                  & --                &   84.4      &   80.9       \\
\textbf{300D SPINN-PI encoders}
                        & 3.7m                  & --                &   89.2      &   \underline{83.2}       \\
\result{\textbf{300D SPINN encoders} }
                        & ??                  & 92?           &   ?    &   82?      \\          
% \result{300D SPINN-PI, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-PI, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
\midrule
\multicolumn{5}{c}{Other previous NN results}\\
\midrule
100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
                        & 252k               & --              &   85.3   &   83.5       \\
300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
                        & 1.9m               & --             &   92.0   &   86.1      \\
300D LSTMN with deep attention fusion \citep{cheng2016long}
                        & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy. \todo{Move attention results to footnote.}} 
\end{table*}

@@

\result{[SB] Compare the types of errors in the different runs and cherry pick interesting examples/tenancies; measure frequency of broken parses.}
